raptr: Representative and Adequate Prioritization Toolkit in R

### 2017-04-06 code/rmarkdown/article.tex
r pkgs

\documentclass[11pt,]{article}
\usepackage{lmodern}
\usepackage{amssymb,amsmath}
\usepackage{ifxetex,ifluatex}
\usepackage{fixltx2e} % provides \textsubscript
\ifnum 0\ifxetex 1\fi\ifluatex 1\fi=0 % if pdftex
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
\else % if luatex or xelatex
  \ifxetex
    \usepackage{mathspec}
    \usepackage{xltxtra,xunicode}
  \else
    \usepackage{fontspec}
  \fi
  \defaultfontfeatures{Mapping=tex-text,Scale=MatchLowercase}
  \newcommand{\euro}{€}
\fi
% use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
% use microtype if available
\IfFileExists{microtype.sty}{%
\usepackage{microtype}
\UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\usepackage[margin=1in]{geometry}
\usepackage{graphicx}
\makeatletter
\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
\makeatother
% Scale images if necessary, so that they will not overflow the page
% margins by default, and it is still possible to overwrite the defaults
% using explicit options in \includegraphics[width, height, ...]{}
\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
\ifxetex
  \usepackage[setpagesize=false, % page size defined by xetex
              unicode=false, % unicode breaks when used with xetex
              xetex]{hyperref}
\else
  \usepackage[unicode=true]{hyperref}
\fi
\hypersetup{breaklinks=true,
            bookmarks=true,
            pdfauthor={},
            pdftitle={rapr: Representative and Adequate Prioritisations in R},
            colorlinks=true,
            citecolor=blue,
            urlcolor=blue,
            linkcolor=magenta,
            pdfborder={0 0 0}}
\urlstyle{same}  % don't use monospace font for urls
\setlength{\parindent}{0pt}
\setlength{\parskip}{6pt plus 2pt minus 1pt}
\setlength{\emergencystretch}{3em}  % prevent overfull lines
\setcounter{secnumdepth}{0}

@@

\setlength{\droptitle}{-2em}
  \title{rapr: Representative and Adequate Prioritisations in R}
  \pretitle{\vspace{\droptitle}\centering\huge}
  \posttitle{\par}
  \author{Jeffrey O. Hanson$^1$, Jonathan R. Rhodes$^2$, Hugh P. Possingham$^1$,
Richard A. Fuller$^1$\\$^1$School of Biological Sciences, The University
of Queensland, Brisbane, QLD, Australia\\$^2$School of Geography,
Planning and Environmental Management, The University of Queensland,
Brisbane, QLD, Australia\\Correspondance should be addressed to
\href{mailto:jeffrey.hanson@uqconnect.edu.au}{jeffrey.hanson@uqconnect.edu.au}}
  \preauthor{\centering\large\emph}
  \postauthor{\par}
  \predate{\centering\large\emph}
  \postdate{\par}
  \date{18 July 2016}

% load packages
\usepackage{amsmath,amsfonts,float,makecell,titletoc,titlesec,tocloft,titling,natbib,pdfpages,lineno,booktabs}
\usepackage[T1]{fontenc}
\usepackage{lmodern}
\usepackage[utf8]{inputenc}
\usepackage[doublespacing]{setspace}

@@

% format toc
%\renewcommand\cftsubsecfont{\normalfont\normalsize\bfseries}
% \renewcommand\cftsubsubsecfont{\normalfont\normalsize\bfseries}
% \renewcommand\cftparafont{\normalfont\normalsize\bfseries}
% \renewcommand\cftsubparafont{\normalfont\normalsize\itshape}

@@

@@

\begin{document}

\maketitle

\begin{abstract}
\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\itemsep1pt\parskip0pt\parsep0pt
\item
  An underlying aim in spatial conservation planning is to maximise the
  long-term persistence of biodiversity. Yet current approaches
  overwhelming focus on ``adequacy'', the concept ensuring a target
  proportion of the distribution of a biodiversity feature (species or
  ecosystem) is reserved. Ultimately, to achieve long-term persistence,
  the processes the sustain biodiversity must also be conserved. Many
  biodiversity processes can be preserved by conserving a representative
  sample of each biodiversity feature. This idea has for decades been
  encapsulated in principle of ``representativeness''. For example,
  representing the full range of genetic diversity or ability to persist
  in a range of climatic conditions might be crucial for the long -term
  persistence of a species in a world where environmental change is
  accelerating. However, fully operationalising the principle of
  repressiveness in spatial conservation planning has proven
  methodologically challenging.
\item
  To address this issue, we developed the
  \texttt{rapr: Representative and Adequate Pioritisations in R}, a
  toolkit to guide reserve selection using explicit targets for
  representing biodiversity processes as well as meeting area protection
  targets. Using a novel formulation of the reserve selection problem,
  users set ``space targets'' to secure a representative sample of each
  species within an attribute space reflecting any biodiversity process
  (eg. an attribute space expressing variation in genetic
  characteristics between individuals in an area, or variation in
  climatic conditions between areas occupied by indivduals). We explored
  the functionality of this \texttt{R} package using a simulation study
  and two case studies. In each study, we generated prioritisations that
  aimed to preserve an amount of habitat for each species using amount
  target--to represent conventional reserve selection methods--and
  compared them with prioritisations generated using both habitat and
  space targets--to represent this method.
\item
  We show that explicitly considering biodiversity processes in reserve
  selection can substantially change the configuration of the resulting
  solution. Simulations suggest that including representativeness in
  conservation planning is particularly important where the biodiversity
  feature has multi-model distributions in an attribute space. Results
  from initial case studies show that explicitly considering space-based
  targets for biodiversity processes can result in prioritisations that
  do not necessarily require much more area than traditional approaches,
  but that could be much more effective in achieving long term
  biodiversity persistence.
\item
  The \texttt{rapr R} package provides a unified framework achieving
  spatial conservation prioritisations that are both adequate and
  representative, resulting in a greater chance of long term
  biodiversity persistence that area-based planning alone.
\end{enumerate}
\end{abstract}

Short running title: Representative and adequate prioritisations
\newline
Abstract word count: 370 / 350\\Total word count: XXXXX / 8000\\Number
of references: 34\\Number of figures: 8\\Number of tables: 0\\Number of
textboxes: 0\\Number of equations: 8\\\newpage

\section{Introduction}\label{introduction}

Perhaps the most fundamental aim of conservation is to maximise the
long-term persistence of biodiversity at organisational levels (McNeely
1994; Margules \& Pressey 2000). To achieve this, conservation actions
must preserve biodiversity patterns (eg. populations, species,
ecosystems), but also crucially the processes that sustain them. One of
the major tangible achievements of modern conservation has been the act
of setting aside areas for preservation (Sanderson \emph{et al.} 2015).
Reserve networks buffer species from gross threatening processes and set
the stage for enhanced management interventions (Gaston \emph{et al.}
\{2008\}). However, the resources available for conservation action are
limited, and so reserve networks must be sited in places that satisfy
conservation objectives for minimum cost (Margules \& Pressey 2000). To
achieve this, reserve selection is often formulated as an optimisation
problem and then solved to identify cost-effective candidate reserve
systems (termed prioritisations; Margules \& Pressey 2000). Many
decision support tools have been developed to help identify
cost-effective solutions (eg. \texttt{ ConsNet}, Ciarleglio \emph{et
al.} 2009;\texttt{ Marxan}, Ball \emph{et al.} 2009;\texttt{ Zonation},
Moilanen 2007). Typically, these tools are used to deliver a
prioritisation that preserves a set of species of interest to the
conservation planner (termed target species or features). Given a set of
species, decision makers can use these tools to preserve biodiversity
patterns by generating a prioritisation that secures an adequate
proportion of each species' range.

However, to achieve effective conservation, reserve networks must do
more than cover enough area. They must capture the processes that
sustain biodiversity over the long term; they must capture both
ecological and evolutionary processes (Crandall \emph{et al.} 2000;
Margules \& Pressey 2000). Ecological processes, such as predator-prey
interactions, pollination, and decomposition, are required for
biodiversity to persist over short time-scales. Typically, they operate
over small geographic domains--with exceptions such as migration and
refugial habitats--and can be preserved using suitably large planning
units (Ciarleglio \emph{et al.} 2009) that each contain a discrete unit
of habitat (Klein \emph{et al.} 2009). On the other hand, evolutionary
processes are required for biodiversity to persist over long
time-scales, and they typically operate over large geographic domains.
Protected areas must represent adaptive evolutionary processes to foster
resilience against environmental change (eg. climate change; Pyke \&
Fischer 2005). Protected areas must also preserve neutral evolutionary
processes (Moritz 1994, 1999, 2002), arising from restriction of gene
flow between populations. They are important for maintaining genetic
diversity, and avoiding inbreeding depression. In recent decades, a
wealth of data relevant to evolutionary processes has become available
to conservation planners (eg. bioclimatic, genetic, and trait data;
Hijmans \emph{et al.} 2005; Raymond \emph{et al.} 2015; Jones \emph{et
al.} 2009). Yet such data are rarely used to guide conservation planning
(Hendry \emph{et al.} 2010). This is perhaps in large part due to that
fact that existing reserve selection tools focus on capturing either
biodiversity patterns or processes--but not both.

Existing conservation planning tools are generally not well-suited for
representing biodiversity processes (but see Faith 2003). To preserve
biodiversity processes, a prioritisation must capture a representative
sample of each species. For instance, to preserve predator-prey
interactions, a prioritisation must preserve individuals from each
predator and prey species in the same area. To preserve adaptive
evolutionary processes, a prioritisation must preserve the adaptive
landscape of each species--populations experiencing different selection
pressures (Moritz 2002). To preserve neutral evolutionary processes, a
prioritisation must secure individuals descended from each of the
genetic lineages that comprise each species (Moritz 1994). Previous
attempts to accommodate such into multi-species conservation planning
have partitioned species' ranges into different groups, for example at
habitat discontinuities (e.g. Carvalho \emph{et al.} 2011) as a
pre-processing step to render a new set of ``pseudo-species''. However,
this approach assumes that the biodiversity processes that operate
across a species' range can be readily split into discrete units. Yet
data on biodiversity processes is often continuous and
hyper-dimensional, and often cannot be reduced to a few categories
without significant information loss (Faith \& Walker 1996).

One of the key issues in reserve selection is the lack of a unifying
decision support tool that can accommodate data on biodiversity patterns
and processes in a multi-species context. To begin to fill this gap, we
present the \texttt{rapr R} package. This \texttt{R} package uses novel
formulations of the reserve selection problem to provide decision makers
with the tools to generate prioritisations that preserve both
biodiversity patterns and processes. We aim to define the concepts
behind the problem formulations. Furthermore, we aim to explore the
functionality of the \texttt{rapr R} package by applying one of the
formulations to a set of simulated species and two case studies.

\section{Methods}\label{methods}

\subsection{PROBLEM FORMULATION}\label{problem-formulation}

Biodiversity features are defined as the entities that the
prioritisation is required to preserve (eg. species, ecosystems).
Spatial attributes are defined as the variation across the species'
range that the prioritisation is required to sample. They can be
intrinsic (eg. genetic or trait variation) or extrinsic (eg.
environmental variation) to the feature. These attributes should be
related to the biodiversity processes that the decision maker aims to
preserve.

Each attribute is conceptualised as a space. This space is termed an
``attribute space''. Each planning unit is thought to occupy a single
point inside each space. For example, a decision maker may require a
prioritisation that represents populations along climatic gradients. To
achieve this, the decision maker might use an ``climatic'' attribute
space with dimensions relating to mean annual temperature ($^{\circ}$C)
and precipitation (mm). Any given combination of temperature and
precipitation may be conceived as a point in this environmental space.
By associating planning units with climatic data, they can be mapped
from geographic space to this environmental attribute space.

Demand points are points that also exist in an attribute space. They are
designated by the decision maker to indicate regions of the attribute
space that the decision maker wishes to represent in the prioritisation
(see below for discussion on how demand points can be generated for
real-world datasets). The amount of variation in the attribute space
that a prioritisation secures is a function of the distance between each
demand point and each selected planning unit in the attribute space. The
shorter the distances between the demand points and the planning units;
the better the prioritisation is at securing the variation in the
spatial attribute. To convert these amounts to a proportion--a
meaningful unit for a decision maker--the distances between the selected
planing units and the demand points are scaled by the distances between
the demand points to the centroid of the demand points. In any attribute
space there may exist points that are impossible (eg. mean annual
rainfall -5 mm), or do not occur in the study area (eg. mean annual
temperature 30$^{\circ}$C in Antarctica). Additionally, there may be
some regions that are desirable for some features and undesirable for
others (eg. conditions known to be outside the physiological tolerance
of certain species). Thus a different set of demand points and weights
are used for each attribute space and each feature. By placing demand
points in desirable regions of an attribute space for a given feature,
the decision maker can ensure that prioritisations secure the feature in
planning units with spatial attributes that are desirable for that
feature.

To illustrate these concepts, consider an example conservation planning
scenario example involving an attribute space and demand points. We wish
to develop a prioritisation for a single species that has four
populations in the study area. However, we can only afford to preserve
three populations. We aim preserve the adaptive landscape of the
species, and to achieve this, we will preserve populations inhabiting
different environmental conditions. To describe environmental variation,
we obtain data on the environmental conditions (rainfall (mm) and
temperature ($^{\circ}$C)) where each population is found. These
environmental data are then used to construct a two-dimensional
environmental attribute space. Next, we generated demand points as
equi-distant points between the range of values where the populations
were found. By comparing the distribution of the demand points to the
distribution of the populations in the attribute space, we can identify
a suitable prioritisation (Fig. 1). We can see that preserving both
populations $A$ and $C$, effectively ``doubles-up'' on the same
environmental characteristics, resulting in considerable redundancy in
the solution. Instead, a more representative sample of the
intra-specific variation could be captured by securing populations $A$,
$B$, and $D$. This example demonstrates how the inclusion of
biodiversity processes can guide the reserve selection process.

The formulations used to express the reserve selection problem in the
\texttt{rapr R} package are based on a combination of the
\texttt{Marxan} reserve selection problem and the uncapacitated facility
location problems (Cornu{é}jols \emph{et al.} 1990). Although the
\texttt{rapr R} package provides two novel formulations, for brevity, we
will define the simpler formulation--referred to as the unreliable
formulation--below and define the more complex version--the reliable
formulation--in the Supporting Information S1. These formulations are
named after the unreliable and reliable facility location formulations
from which they are based upon (Cui \emph{et al.} 2010). The key
difference between these two formulations is that the reliable
formulation explicitly considers the probability that planning units are
occupied when calculating the proportion of an attribute space sampled
in a solution. All mathematical terms defined hereafter are described in
Table S1. For convenience, the cardinality of sets will be denoted using
the same symbol used to denote the variable.

Define $F$ to be the set of features one wishes to conserve (indexed by
$f$). Let $J$ be a set of planning units (indexed by $j$). Also, let
$A_j$ denote the area, and $C_j$ denote the cost of preserving planning
unit $j \in J$. To assess the extent to which each feature is secured in
a given prioritisation, let $q_{fj}$ denote the probability of feature
$f$ occupying planning unit $j$. The level of fragmentation associated
with a prioritisation is parametrised as the net exposed boundary
length. Let the shared edges between each planning unit $j \in J$ and
$k \in J$ be $e_{jk}$.

Let $S$ denote a set of attribute spaces (indexed by $s$). Each
$j \in J$ is associated with spatially explicit data that represent
coordinates for each attribute space $s \in S$. Let $I_{fsi}$ denote a
set of demand points (indexed by $i$) for each feature $f \in F$ and
each attribute space $s \in S$. Let $\lambda_{fsi}$ denote the weighting
for each demand point $i \in I$, $f \in F$ and $s \in S$. Let $d_{fsij}$
denote the distance between each demand point $i \in I$ and each
planning unit $j \in J$ for each feature $f \in F$ and attribute space
$s \in S$. To describe the inherent variation in the distribution of
demand points for feature $f$ and space $s$, let $\delta_{fsi}$ denote
the distance between each demand point $i \in I$ and the centroid of the
demand points. Demand points with greater weight $\lambda_{fsi}$ are
more important, and the optimal solution will be likely to select
planning units close to highly weighted demand points. As a consequence,
the decision maker will need to choose an appropriate weighting for each
demand point.

Targets are used to ensure that prioritisations adequately preserve each
species. Amount-based targets are used to ensure that the total amount
of habitat preserved is sufficient. Let $T_f$ denote the expected amount
of area that needs to be preserved for each feature $f \in F$.
Space-based targets ensure that a sufficient proportion of the
intra-specific variation is secured. Let $\tau_{fs}$ denote the
space-based targets for feature $f \in F$ and attribute space $a \in A$.
For convenience, these both types of targets are expressed as
proportions in the \texttt{R} package.

The control variables for the unreliable formulation are the $B$,
$T_{s}$, and $\tau_{sa}$ variables.

\begin{align*}
T_s &= \text{amount target for feature $f$} \tag*{eqn 1a}\\
%
\tau_{sa} &= \text{representation target for feature $f$ in attribute space $a$} \tag*{eqn 1b}\\
%
B &= \text{boundary length modifier (BLM): penalise fragmented solutions} \tag*{eqn 1c}\\
\end{align*}

The decision variables are the $X_j$ and $Y_{fsij}$ variables.

\begin{align*}
X_j
    &= \begin{cases}
        1, & \parbox{25em}{if planning unit $j$ is selected for conservation action} \tag*{eqn 2a} \\
        0, & \parbox{25em}{otherwise} \\
    \end{cases} \\
%
Y_{fsij} &= \begin{cases}
        1, & \parbox{25em}{if demand point $i$ is assigned to planning unit $j$ for feature $f$ in space $s$. } \tag*{eqn 2b} \\
        0, & \parbox{25em}{otherwise} \\
    \end{cases} \\
\end{align*}

Each demand point $i \in I$ for feature $f \in F$ and space $s \in S$ is
assigned to a selected planning unit $J$ where $X_j = 1$. The weighted
distance between the demand point and its assigned planning unit
$\lambda_{fsi} d_{fsij}$ is used to assess how well the demand point is
represented in a given solution. Generally, demand points are assigned
to the closest selected planning units (unless particularly low
space-based targets are used).

The unreliable formulation (URAP) is a defined as a multi-objective
optimisation problem.

\begin{align*}
& \text{(URAP)} & \text{Min } & \sum_{j=0}^{J-1} \left( X_j C_j \right) + \sum_{j=0}^{J-1} \sum_{k=j}^{J-1} X_j \left( 1-X_k \right) \left( B e_{jk} \right) + & & \tag*{eqn 3a} \\
%
& & \text{s.t. } & \sum_{j=0}^{J-1} A_j q_{fj} \geq T_{f} & \forall & 0 \leq f \leq F-1 \tag*{eqn 3b}\\
%
& & & 1 - \frac{\sum_{i=0}^{I-1} \sum_{j=0}^{J-1} \lambda_{fsi} {d_{fsij}}^{2} Y_{fsij}}{\sum_{i=0}^{I-1} \lambda_{fsi} {\delta_{fsi}}^{2}} \geq \tau_{fs} & \forall & 0 \leq f \leq F-1, \tag*{eqn 3c}\\
& & & & & 0 \leq s \leq S-1\\
%
& & & \sum_{j=0}^{J-1} Y_{fsij} = 1 & \forall & 0 \leq f \leq F-1, \tag*{eqn 3d}\\
& & & & & 0 \leq s \leq S-1, \\
& & & & & 0 \leq i \leq I-1\\
%
& & & Y_{fsij} \leq X_j & \forall & 0 \leq f \leq F-1, \tag*{eqn 3e}\\
& & & & & 0 \leq s \leq S-1, \\
& & & & & 0 \leq i \leq I-1,\\
& & & & & 0 \leq j \leq J-1\\
%
& & & X_j, Y_{fsij} \in {0,1} & \forall & 0 \leq f \leq F-1, \tag*{eqn 3f}\\
& & & & & 0 \leq s \leq S-1,\\
& & & & & 0 \leq i \leq I-1\\
%
\end{align*}

The objective function (eqn 3a) determines the utility of a given
prioritisation: a combination of the total cost of a prioritisation and
how fragmented it is. Constraints (eqn 3b) ensure that all the
amount-based targets are met. Constraints (eqn 3c) ensure that all the
space-based targets are met for each feature and each attribute space.
For each feature and attribute space, the total weighted distance
between the demand points and their closest selected planning units is
calculated ($\lambda_{fsi} d_{fsij} Y_{fsij}$). This total weighted
distance is then scaled by the inherent variation in the demand points
($\lambda_{fsi} \delta_{fsi}$). The resulting fraction yields a
proportion conceptually similar to the $R^2$ statistic used in $k$-means
clustering analysis. The constraints ensure that this proportion must be
equal to or greater than the space-based target. Constraints (eqn 3d)
ensure that only one planning unit is assigned to each demand point.
Constraints (eqn 3e) ensure that demand points are only assigned to
selected planning units. Constraints (eqn 3f) ensure that the $X$ and
$Y$ variables are binary.

\subsection{OPTIMISATION}\label{optimisation}

The unreliable formulation is non-linear. However, the non-linear
components can be linearised using existing techniques. The expression
$X_j X_k$ in (eqn 3a) can be linearised using methods described by Beyer
et al. (2016). Linearised versions of the problems can be solved using
commercial exact algorithm solvers. The \texttt{rapr} R package provides
functions to express conservation planning data as optimisation problems
using linearised versions of the unreliable and reliable formulations.
These optimisation problems can then be solved to generate
prioritisations using the commercial \texttt{Gurobi} software suite
(\url{http://www.gurobi.com}). Presently, academics can obtain a license
at no cost from the Gurobi website. After installing the \texttt{Gurobi}
software suite, users will need to install the \texttt{Gurobi R}
package.

\section{Examples}\label{examples}

To understand the behaviour of the unreliable problem and showcase its
value, we conducted a simulation study and two case studies. These
studies involved generating solutions using only amount targets to
represent prioritisations generated using conventional methods (eg.
\texttt{Marxan}), and solutions using amount and space targets using the
unreliable formulation. By comparing these solutions, we can guage the
benefits of explicitly including space targets in reserve selection. All
analyses were performed in \texttt{R} (version 3.3.0; {R Core Team}
2016).

\subsection{SIMULATION STUDY}\label{simulation-study}

\subsubsection{Methods}\label{methods-1}

We simulated a hypothetical study area with planning units arranged in
$10 \times 10$ square grid (Fig. 2). We then simulated three species
across this study area. The first species was simulated to represent a
hyper-generalist--occurring in all planning units with equal probability
(Fig. 2a). This species' distribution was based on a uniform
distribution (eqn 8a). The second species was simulated to represent a
species with an idealised distribution (Fig. 2b). It has a core range
area, and is less likely to be found in areas that are distant from the
core area. This species' distribution was simulated using the
probability density function of a single multivariate normal
distribution (eqn 8c). The third species represents a species with two
distinct populations (Fig. 2c). This species' distribution was simulated
to depict a bimodal distribution, based on the combination of the
probability density functions of two multivariate normal distributions
(eqn 8d). The probability that each species inhabited a given planning
unit was calculated using the $X, Y$ coordinates of the planning unit
and eqns 8a--8d. We used a geographic attribute space to showcase the
behavior of the problem. For each species, demand points were generated
by calculating the centroids of each planning units and their weights
were set as the probability that the planning units were occupied.

\begin{align*}
\\
P \left( \text{uniformly distributed species} | \left( x, y \right) \right) &= 0.1 \tag*{eqn 8a} \\
%
f \left(z, \mu, \Sigma \right) &= \left(2 \pi \right) ^{-1} | \Sigma | e ^{- \frac{1}{2} \left(z - \mu \right)' \Sigma ^{-1} \left(z - \mu \right)} \tag*{eqn 8b} \\
%
P \left( \text{normally distributed species} | \left( x , y \right) \right) &= \frac{f \left( \left[ \begin{smallmatrix} x \\ y \end{smallmatrix} \right] \, \left[ \begin{smallmatrix} 0.0 \\ 0.0 \end{smallmatrix} \right] \, \left[ \begin{smallmatrix} 12.58&0 \\ 0&12.5 \end{smallmatrix} \right] \right)}{2} \tag*{eqn 8c} \\
%
P \left( \text{bimodally distributed species} | \left( x,y \right) \right) &= \text{Max} \begin{cases}
f \left(\left[ \begin{smallmatrix} x \\ y \end{smallmatrix} \right] \, \left[ \begin{smallmatrix} -3.75 \\ -3.75 \end{smallmatrix} \right] \, \left[ \begin{smallmatrix} 10&0 \\ 0&10 \end{smallmatrix} \right] \right), \\ \frac{f \left( \left[ \begin{smallmatrix} x \\ y \end{smallmatrix} \right] \, \left[ \begin{smallmatrix} 3.75 \\ 3.75 \end{smallmatrix} \right] \, \left[ \begin{smallmatrix} 8&0 \\ 0&8 \end{smallmatrix} \right] \right)} {2} \\
\end{cases} \tag*{eqn 8d} \\
%
\end{align*}

We generated four solutions for each species. First to represent
solutions generated using conventional conservation planning methods, we
generated solutions using only 20 \% amount targets. Second to show how
the addition of geographic targets can affect a prioritisation, we
generated solutions using 20 \% amount targets and 75 \% geographic
targets. Third to represent solutions using conventional planning
methods that penalise for fragmentation, we generated solutions using 20
\% amount targets and a boundary length modifier of 1. Fourth to
illustrate the combined affect using geographic targets and penalising
fragmentation, we generated solutions using using 20 \% amount targets
and 75 \% geographic targets and a boundary length modifier of 1.

\subsubsection{Results}\label{results}

The uniform species was simulated to occur with a constant probability
of occupancy across the study area (Fig. 2a). The solution generated
using 20 \% amount-based targets (Fig. 3a) selected 20 planning units
near the southern end of study area. This configuration is an artifact
of the method used to solve this particular instance of the reserve
selection problem. In fact, for this species, all prioritisations
containing 20 planning units are optimal when considering only 20 \%
amount targets. In the absence of criteria to guarantee
representativeness, reserve selection methods may or may not return
solutions that secure a representative sample of the features. In this
particular case, the solution does not secure a representative sample of
the species' range (-23.64 \% sampled).

The addition of a 75 \% geographic target resulted in a solution that
secured a representative sample of the uniform species' range (90.67 \%
sampled; Fig. 3b). Since all planning units have equal cost and an equal
chance of being occupied, this solution has the same number of planning
units as the solution generated using only amount targets (cf.~Fig 3a).
Although the use of amount and space based targets has addressed
adequacy and representativeness objectives (respectively), they have
resulted in a highly fragmented solution.

The addition of a positive boundary length modifier (BLM) parameter
resulted in a well-connected solution that secured an adequate
proportion (20 \% ) and representative sample (75.76 \% sampled) of the
species geographic distribution (Fig. 3d). This solution contains 20
planning units--a few more than the previously discussed solutions--to
ensure that the solution secures a representative proportion of the
species geographic distribution in a configuration that is not highly
fragmented. This result suggests that the combination of space targets
and boundary length modifiers may yield solutions that contain more
planning units, since the solution generated using amount targets and a
boundary length modifier (Fig. 3c) contained the sample number of
planning units as the solution with just the amount targets (Fig. 3a).
Overall, these results show that under the simplest of conditions, the
reliance on just amount targets can cause an under-specified reserve
selection problem that is unlikely to return a suitable solution for
implementation.

The normally distributed species was simulated to contain a single core
area where individuals are most prevalent and marginal areas where
individuals are less likely to occur (Fig. 2b). The solution generated
using 20 \% amount targets contained 10 planning units and concentrated
conservation efforts in the core area (Fig. 3e). Since all planning
units have equal costs and areas, this solution contains the planning
units with the highest probabilities of occupancy. Whilst this solution
satisfies the adequacy objective for a prioritisation in a
cost-effective manner--it fails to fulfill the representative objective
for a prioritisation (59.09 \% distribution sampled).

The solution generated using 20 \% amount targets and 75 \% geographic
targets resulted in a solution that is both adequate and representative
in terms of the uniform species distribution (Fig. 3f). This solution
used 11 planning units to secure 20.02 \% and sample 76.31 \% of the
normal species' distribution. By using amount and geographic targets in
the reserve selection problem, we have obtained a solution that
concentrates conservation effort in the range core and also the range
margin. However, similar to the corresponding solution for the uniform
species (Fig. 3b), this solution is highly fragmented.

By using 20 \% amount targets and 75 \% geographic targets and a
boundary length modifier parameter ($BLM = 1$), we obtained a solution
that fulfills adequacy, representativeness and connectivity objectives
(Fig. 3h). This solution contains 12 planning units. Similar to the
corresponding solution for the uniform species (Fig. 3d), this solution
contains more planning units than any other solutions for this species
(cf.~Figs. 3e--3g). The results for the normally distributed species
suggest that the space targets can result in solutions that secure a
more representative sample of the species'--even for species without
significant structure.

The bimodally distributed species was simulated to represent a species
with a highly structured distribution, and it contains two distinct
populations (Fig. 2c). The solution generated using just 20 \% amount
targets assigned conservation effort to just one of the two populations
(Fig. 3i). While this solution secured an adequate proportion of the
species' distribution (20.18 \% secured), it did not sample a
representative proportion of the range (12 \% sampled). The addition of
boundary length modifiers to just amount targets resulted in a solution
which sampled even less of the species distribution (25.19 \% sampled;
Fig. 3k). However, the addition of geographic space targets resolved
these issues.

>>>>

\documentclass[11pt,]{article}
\usepackage{lmodern}
\usepackage{amssymb,amsmath}
\usepackage{ifxetex,ifluatex}
\usepackage{fixltx2e} % provides \textsubscript
\ifnum 0\ifxetex 1\fi\ifluatex 1\fi=0 % if pdftex
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
\else % if luatex or xelatex
  \ifxetex
    \usepackage{mathspec}
  \else
    \usepackage{fontspec}
  \fi
  \defaultfontfeatures{Ligatures=TeX,Scale=MatchLowercase}
\fi
% use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
% use microtype if available
\IfFileExists{microtype.sty}{%
\usepackage{microtype}
\UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\usepackage[margin=1in]{geometry}
\usepackage{hyperref}
\hypersetup{unicode=true,
            pdftitle={raptr: Representative and Adequate Prioritization Toolkit in R},
            pdfborder={0 0 0},
            breaklinks=true}
\urlstyle{same}  % don't use monospace font for urls
\usepackage{graphicx,grffile}
\makeatletter
\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
\makeatother
% Scale images if necessary, so that they will not overflow the page
% margins by default, and it is still possible to overwrite the defaults
% using explicit options in \includegraphics[width, height, ...]{}
\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
\IfFileExists{parskip.sty}{%
\usepackage{parskip}
}{% else
\setlength{\parindent}{0pt}
\setlength{\parskip}{6pt plus 2pt minus 1pt}
}
\setlength{\emergencystretch}{3em}  % prevent overfull lines
\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
\setcounter{secnumdepth}{0}

@@

\setlength{\droptitle}{-2em}
  \title{raptr: Representative and Adequate Prioritization Toolkit in R}
  \pretitle{\vspace{\droptitle}\centering\huge}
  \posttitle{\par}
  \author{Jeffrey O. Hanson\(^1\), Jonathan R. Rhodes\(^2\), Hugh P.
Possingham\(^{1,3}\), Richard A. Fuller\(^1\)\\
\(^1\)School of Biological Sciences, The University of Queensland,
Brisbane, QLD, Australia\\
\(^2\)School of Earth and Environmental Sciences, The University of
Queensland, Brisbane, QLD, Australia\\
\(^3\)The Nature Conservancy, South Brisbane, QLD, Australia\\
Correspondence should be addressed to
\href{mailto:jeffrey.hanson@uqconnect.edu.au}{\nolinkurl{jeffrey.hanson@uqconnect.edu.au}}}
  \preauthor{\centering\large\emph}
  \postauthor{\par}
  \predate{\centering\large\emph}
  \postdate{\par}
  \date{06 April 2017}

% load packages
\usepackage{amsmath,amsfonts,float,makecell,titletoc,titlesec,tocloft,lineno,booktabs,subfiles}
\usepackage[T1]{fontenc}
\usepackage{lmodern}
\usepackage[utf8]{inputenc}
\usepackage[doublespacing]{setspace}

@@

@@

% allow breaks in equations
\allowdisplaybreaks

@@

\begin{document}
\maketitle

Abstract word count: 347 / 350\\
Total word count: XXXX / 7000\\
Number of references: 42\\
Number of figures: 8\\
Number of tables: 0\\
Number of textboxes: 0\\
Number of equations: 4\\
Short running title: Short abstract: Prioritize biodiversity processes
and patterns using raptr \url{https://github.com/jeffreyhanson/raptr}\\
Keywords: conservation, biodiversity, mixed integer linear programming,
optimization, protected areas

\clearpage

\section{Summary}\label{summary}

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\item
  An underlying aim in conservation planning is to maximize the
  long-term persistence of biodiversity. To fulfill this aim, the
  ecological and evolutionary processes that sustain biodiversity must
  be preserved. One way to conserve such processes at the feature level
  (eg. species, ecosystem) is to preserve a sample of the feature (eg.
  individuals, areas) that is representative in terms of the intrinsic
  or extrinsic physical attributes that underpin the process of
  interest. For example, by conserving a sample of populations with
  local adaptations that is representative in terms of the full
  variation found in the species--physical attributes associated with
  adaptation--protected areas can foster adaptive processes by ensuring
  these adaptations are not lost. Despite this, current reserve
  selection methods overwhelmingly focus on securing an adequate amount
  of area or habitat for each feature. Little attention has been
  directed towards capturing a representative sample of each feature.
\item
  To address this issue, we developed the \texttt{raptr R} package to
  help guide reserve selection. Users set ``amount targets''--similar to
  conventional methods--to ensure that solutions secure a sufficient
  proportion of area or habitat for each feature. Additionally, users
  set ``space targets'' to secure a representative sample of variation
  in ecologically or evolutionary relevant attributes (eg. environmental
  or genetic variation). We demonstrate the functionality of this
  package using simulations and two case studies. In these studies, we
  generated solutions using amount targets--representing conventional
  methods--and compared them with solutions generated using amount and
  space targets.
\item
  Our results demonstrate that different solutions emerge when targeting
  a representative sample of each feature. We show that using these
  targets is important for features that have multimodal distributions
  in the process-related attributes (eg. species with multimodal
  niches). We also found that solutions could conserve a far more
  representative sample with only a slight increase in reserve size.
\item
  The \texttt{raptr R} package provides a toolkit for making
  prioritizations that secure an adequate and representative sample of
  each feature. By using solutions that secure a representative sample
  of each feature, prioritizations may have a greater chance of
  achieving long-term biodiversity persistence.
\end{enumerate}

\clearpage
\input{./text.tex}

\clearpage
\input{./figures.tex}
